---
title: "factorial design-析因设计"
author: "ManBuXiaoCun"
date: '2022-04-24'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
All the code and text are from the book\
and From Data to Decisions: Measurement, Uncertainty, Analysis, and Modeling
Instructor: **Chris Mack**\
**Design and Analysis of Experiments with R**(2015)
Book by **John Lawson**

Chris Mack Course website: http://www.lithoguru.com/scientist/statistics/course.html

### factorial design 析因设计

- 完全析因设计（完全随机设计）
In a factorial design the cells consist of all possible combinations
of the levels of the factors under study. Factorial designs accentuate the factor
eﬀects, allow for estimation of interdependency of eﬀects (or interactions),
and are the ﬁrst technique in the category of what is called treatment design.
(cite from **Design and Analysis of Experiments with R**(2015)
Book by **John Lawson**)

Two-Factor Factorial Plan in R `r expand.grid`
```{r}
D <- expand.grid( BW = c(3.25, 3.75, 4.25), WL = c(4, 5, 6) )
D  # original design when no replicate
D <- rbind(D,D) # duplicates for experiments

set.seed(2591)
D <- D[order(sample(1:18)), ] # randomize the order of D
CopterDes <- D[ c( "BW", "WL" )]
CopterDes
write.csv(CopterDes, file = "CopterDes.csv", row.names = FALSE) # save design to csv file
```


More than two factors Factorial Design
Package **AlgDesign** **gen.factorial()**

```{r}
library(AlgDesign)
Design_34 <- gen.factorial(levels = 3, nVars = 4, varNames = c(paste0("factor",1:4))) 
Design_34                           
# 4因素3水平，因素名字factor1,factor2,factor3,factor4

#save into csv file
write.csv(Design_34, file = "Design_34.csv", row.names = FALSE)
```

Analysis of two factors Factorial Design
```{r}
library(daewr)
data(COdata)
head(COdata)
model1 <- aov( CO ~ Eth * Ratio, data = COdata )
summary(model1)
```

AlgDesign for Factorial Design
```{r}
library(AlgDesign)
# As an example, we'll create full factorial design with three factors (F1, F2, F3) and
# with 3, 2, and 3 levels, respectively

# 3, 2, 3 levels, 3 factors
Design323L3F <- gen.factorial(levels = c(3,2,3), nVars = 3, center=TRUE, varNames=c("F1", "F2", "F3"))
# When center = TRUE, the design is symmetric and the three levels will be signified by -1, 0, and 1

# The common 2^k design (two levels for each factor):
dat = gen.factorial(levels = 2, nVars = 3, varNames=c("F1", "F2", "F3"))
dat
# the output is a data frame containing the factorial design

#
write.csv(Design323L3F, file = "Design323L3F.csv", row.names = FALSE)

# 写入指定文件夹 write it into data folder
write.csv(Design323L3F, file = here::here("data","Design323L3F.csv"), row.names = FALSE)
```

Analysis of a factorial design
```{r}
library(tidyverse)
library(readr)
yield <- read_csv( here::here("FactorialDesign.csv") )
yield
# to remove yield column 1
yield <- yield[-1]
plot(yield)
```

```{r}
library(lattice)
xyplot(Yield ~ Temp | Catalyst, data = yield,
       panel = function(x, y, ...)
       {
         panel.xyplot(x, y, ...)
         panel.lmline(x, y, ...)
       }
)
xyplot(Yield ~ Conc | Catalyst, data = yield,
       panel = function(x, y, ...)
       {
         panel.xyplot(x, y, ...)
         panel.lmline(x, y, ...)
       }
)
```

Explore the "Yield" data from "FactorialDesign.csv"
```{r}
# let's model the main effects
model1 = lm(Yield ~ ., data = yield)
summary(model1)
```

```{r}
# Now we'll add two-factor interactions
model2 = lm(Yield ~ .^2, data = yield)
summary(model2)
```

```{r}
#compare the two models
anova(model1, model2)
```

